---
title: "BeattieAlexandra_final_draft"
author: "Alexandra Beattie"
date: "12/4/2017"
output: word_document
---

```{r echo=FALSE}


library(dplyr)
library(acs)
library(stats)
library(base)
library(fpc)
library(ggplot2)    #for graphics and the dataset for this example session
library(cluster)    #provides more cluster algorithms than base R (e.g. PAM)
library(useful)     #provides a plot function for clusters and "FitKMeans" and "PlotHartigan" functions
library(NbClust)    #provides tons of tools for identifying the "right" number of clusters
library(rgl)        #for 3D rotating plots
library(dbscan)

##Pull in census data to create census based variables

states = geo.make(state='*')

createRace <- function(endYear) {
  b03002 <- acs.fetch(geography=states, table.number="B02001", endyear=endYear, span=1)
  
  stateRace <- data.frame(cbind(data.frame(b03002@geography), data.frame(b03002@estimate))) %>%
    rowwise() %>% summarize(
      fips=paste0(state),
      percent_white=B02001_002/B02001_001,
      percent_black=B02001_003/B02001_001,
      percent_native=B02001_004/B02001_001,
      percent_asian=B02001_005/B02001_001,
      percent_pacific = B02001_006/B02001_001,
      percent_other = B02001_007/B02001_001,
      percent_mixed = B02001_008/B02001_001
    )
  return(stateRace)
}
race2012 <- write.csv("Users/lex/race2012.csv")
race2012 <- createRace('2012')
race2013 <- createRace('2013')
race2014 <- createRace('2014')
race2015 <- createRace('2015')

createAge <- function(endYear) {
  b01002 <- acs.fetch(geography=states, table.number="B01002", endyear=endYear, span=1)
  
  stateAge <- data.frame(cbind(data.frame(b01002@geography), data.frame(b01002@estimate))) %>%
    rowwise() %>% summarize(
      fips=paste0(state),
      median_age=B01002_001
    )
  return(stateAge)
}

age2012 <- createAge('2012')
age2013 <- createAge('2013')
age2014 <- createAge('2014')
age2015 <- createAge('2015')

createPop <- function(endYear) {
  b01003 <- acs.fetch(geography=states, table.number="B01003", endyear=endYear, span=1)
  
  statePop <- data.frame(cbind(data.frame(b01003@geography), data.frame(b01003@estimate))) %>%
    rowwise() %>% summarize(
      fips=paste0(state),
      population=B01003_001
    )
  return(statePop)
}

pop2012 <- createPop('2012')
pop2013 <- createPop('2013')
pop2014 <- createPop('2014')
pop2015 <- createPop('2015')

createAssistance <- function(endYear) {
  b19067 <- acs.fetch(geography=states, table.number="B19067", endyear=endYear, span=1)
  
  stateAssistance <- data.frame(cbind(data.frame(b19067@geography), data.frame(b19067@estimate))) %>%
    rowwise() %>% summarize(
      fips=paste0(state),
      publicAssistance=B19067_001
    )
  return(stateAssistance)
}

assistance2012 <- createAssistance('2012')
assistance2013 <- createAssistance('2013')
assistance2014 <- createAssistance('2014')
assistance2015 <- createAssistance('2015')

createPoverty <- function(endYear) {
  b17001 <- acs.fetch(geography=states, table.number="B17001", 
                      variable = "B17001_001", endyear=endYear, span=1)
  
  statePoverty <- data.frame(cbind(data.frame(b17001@geography), data.frame(b17001@estimate))) %>%
    rowwise() %>% summarize(
      fips=paste0(state),
      povertyIncome=B17001_001
    )
  return(statePoverty)
}

poverty2012 <- createPoverty('2012')
poverty2013 <- createPoverty('2013')
poverty2014 <- createPoverty('2014')
poverty2015 <- createPoverty('2015')

createIncome <- function(endYear) {
  b19013 <- acs.fetch(geography=states, table.number="B19013", endyear=endYear, span=1)
  
  stateMedianIncome <- data.frame(cbind(data.frame(b19013@geography), data.frame(b19013@estimate))) %>%
    rowwise() %>% summarize(
      fips=paste0(state),
      medianIncome=B19013_001
    )
  return(stateMedianIncome)
}

income2012 <- createIncome('2012')
income2013 <- createIncome('2013')
income2014 <- createIncome('2014')
income2015 <- createIncome('2015')


createUnemploy <- function(endYear) {
  b23025 <- acs.fetch(geography=states, table.number="B23025", endyear=endYear, span=1)
  
  stateUnemployement <- data.frame(cbind(data.frame(b23025@geography), data.frame(b23025@estimate)))%>%
    rowwise() %>% summarize(
      fips=paste0(state),
      employed=B23025_004/B23025_003,
      unemployed=B23025_005/B23025_003
    )
  return(stateUnemployement)
}

unemploy2012 <- createUnemploy('2012')
unemploy2013 <- createUnemploy('2013')
unemploy2014 <- createUnemploy('2014')
unemploy2015 <- createUnemploy('2015')


createEd <- function(endYear) {
  c15003 <- acs.fetch(geography=states, table.number="C15003", endyear=endYear, span=1)
  
  stateEducation <- data.frame(cbind(data.frame(c15003@geography), data.frame(c15003@estimate)))%>%
    rowwise() %>% summarize(
      fips=paste0(state),
      none=C15003_002/C15003_001,
      elementary=(C15003_003+C15003_004)/C15003_001,
      junior =(C15003_005+C15003_006)/C15003_001,
      hsNoDiploma = (C15003_007+C15003_008+C15003_009)/C15003_001,
      hsDiploma = C15003_010/C15003_001,
      ged = C15003_011/C15003_001,
      someCollege = (C15003_012+C15003_013)/C15003_001,
      associates = C15003_014/C15003_001,
      bachelors = C15003_015/C15003_001,
      masters = C15003_016/C15003_001,
      professionalDegree = C15003_017/C15003_001,
      doctorate = C15003_018/C15003_001
    )
  return(stateEducation)
}

ed2012 <- createEd('2012')
ed2013 <- createEd('2013')
ed2014 <- createEd('2014')
ed2015 <- createEd('2015')


#USE FOR avg days to wait, main substance admitted and frequency levels for users

#USE FOR # of discharges, avg length of stay, reason for discharge, and priors

#https://www.cdc.gov/drugoverdose/data/statedeaths.html
#for state related deaths

census2015 <- merge(assistance2015, ed2015)
census2015 <- merge(census2015, race2015)
census2015 <- merge(census2015, income2015)
census2015 <- merge(census2015, pop2015)
census2015 <- merge(census2015, unemploy2015)

census2014 <- merge(assistance2014, ed2014)
census2014 <- merge(census2014, race2014)
census2014 <- merge(census2014, income2014)
census2014 <- merge(census2014, pop2014)
census2014 <- merge(census2014, unemploy2014)

census2013 <- merge(assistance2013, ed2013)
census2013 <- merge(census2013, race2013)
census2013 <- merge(census2013, income2013)
census2013 <- merge(census2013, pop2013)
census2013 <- merge(census2013, unemploy2013)

census2012 <- merge(assistance2012, ed2012)
census2012 <- merge(census2012, race2012)
census2012 <- merge(census2012, income2012)
census2012 <- merge(census2012, pop2012)
census2012 <- merge(census2012, unemploy2012)

stateCensusList <- c("AL", "AK", "AZ", "AR", "CA", "CO", "CT", "DE", "DC", "FL", "GA",
                     "HI", "ID", "IL", "IN", "IA", "KS", "KY", "LA", "ME", "MD", "MA",
                     "MI", "MN", "MS", "MO", "MT", "NE", "NV", "NH", "NJ", "NM", "NY",
                     "NC", "ND", "OH", "OK", "OR", "PA", "RI", "SC", "SD", "TN", "TX",
                     "UT", "VT", "VA", "WA", "WV", "WI", "WY", "PR")

census2012 <- cbind(state = stateCensusList, census2012)
census2013 <- cbind(state = stateCensusList, census2013)
census2014 <- cbind(state = stateCensusList, census2014)
census2015 <- cbind(state = stateCensusList, census2015)

##Substance abuse facilities data
load("/Users/lex/Downloads/N-SSATS-2015-DS0001-bndl-data-r 2/N-SSATS-2015-DS0001-data/N-SSATS-2015-DS0001-data-r.rda")
load("/Users/lex/Downloads/N-SSATS-2014-DS0001-bndl-data-r/N-SSATS-2014-DS0001-data/N-SSATS-2014-DS0001-data-r.rda")
load("/Users/lex/Downloads/N-SSATS-2013-DS0001-bndl-data-r/N-SSATS-2013-DS0001-data/N-SSATS-2013-DS0001-data-r.rda")
load("/Users/lex/Downloads/N-SSATS-2012-DS0001-bndl-data-r/N-SSATS-2012-DS0001-data/N-SSATS-2012-DS0001-data-r.rda")

nsafd2015 <- nssats2015_puf
nsafd2014 <- nssatpuf_2014
nsafd2013 <- nssatpuf_2013
nsafd2012 <- nssatpuf_2012

createNSAFD <- function(dataFrame) {
  dataFrame[is.na(dataFrame)] <- 0
  data <- dataFrame[complete.cases(dataFrame),]
  data <- data[,-1]
  data[] <- lapply(data, factor)
  
  dataMatrix <- model.matrix(STATE~.,data)[,-1]
  
  dataDF <- data.frame(dataMatrix)
  
  dataDF$state <- data$STATE
  dataDF$availableService <- 1
  
  summedFin <- dataDF %>% group_by(state) %>% summarise_all(funs(sum))
  
  return(summedFin)
}

nsafd2015 <- createNSAFD(nsafd2015)
nsafd2015$state <- as.factor(nsafd2015$state)
nsafd2015 <- nsafd2015[-c(11, 13, 42, 43, 51), ]
nsafd2015$state <- c("AK", "AL", "AZ", "AR", "CA", "CO", "CT", "DC", "DE", "FL", "GA", "HI",
                     "IA", "ID", "IL", "IN", "KS", "KY", "LA", "MA", "MD", "ME", "MI", "MN", 
                     "MO", "MS", "MT", "NC", "ND", "NE", "NH", "NJ", "NM", "NV", "NY", "OH",
                     "OK", "OR", "PA", "RI", "SC", "SD", "TN", "TX", "UT", "VA", "VT", "WA",
                     "WI", "WV", "WY")
nsafd2014 <- createNSAFD(nsafd2014)
nsafd2013 <- createNSAFD(nsafd2013)
nsafd2012 <- createNSAFD(nsafd2012)

load("/Users/lex/Downloads/TEDSA_2015_PUF")
load("/Users/lex/Downloads/TEDSA_2014_PUF")
load("/Users/lex/Downloads/TEDSA_2013_PUF")
load("/Users/lex/Downloads/TEDSA_2012_PUF")

createTEDSA <- function(dataFrame) {
  data <- dataFrame[, c("STFIPS", "NOPRIOR", "SUB1", "FREQ1")]
  data$NOPRIOR <- as.factor(data$NOPRIOR)
  data$SUB1 <- as.factor(data$SUB1)
  data$FREQ1 <- as.factor(data$FREQ1)
  data <- data[complete.cases(data),]
  dataMatrix <- model.matrix(STFIPS~.,data)[,-1]
  dataDF <- data.frame(dataMatrix)
  dataDF$state <- data$STFIPS
  dataDF$admit <- 1
  summedFin <- dataDF %>% group_by(state) %>% summarise_all(funs(sum))
  return(summedFin)
}
stateCodes <- read.csv("/Users/lex/Documents/stateCodes.csv")


tedsA2014 <- createTEDSA(tedsa_2014)
tedsA2014 <- merge(stateCodes, tedsA2014)
tedsA2015 <- createTEDSA(tedsa_2015)
tedsA2015 <- merge(stateCodes, tedsA2015)
tedsA2013 <- createTEDSA(tedsa_2013)
tedsA2013 <- merge(stateCodes, tedsA2013)
tedsA2012 <- createTEDSA(tedsa_2012)
tedsA2012 <- merge(stateCodes, tedsA2012)

tedsA2012 <- tedsA2012[,-1]
tedsA2013 <- tedsA2013[,-1]
tedsA2014 <- tedsA2014[,-1]
tedsA2015 <- tedsA2015[,-1]

tedsA2012$state <- tedsA2012$stateAbrev
tedsA2013$state <- tedsA2013$stateAbrev
tedsA2014$state <- tedsA2014$stateAbrev
tedsA2015$state <- tedsA2015$stateAbrev

tedsA2012 <- tedsA2012[,-1]
tedsA2013 <- tedsA2013[,-1]
tedsA2014 <- tedsA2014[,-1]
tedsA2015 <- tedsA2015[,-1]

load("/Users/lex/Downloads/teds_d_2014")
load("/Users/lex/Downloads/teds_d_2013")
load("/Users/lex/Downloads/teds_d_2012")
```
```{r}



createTEDSD <- function(dataFrame) {
  data <- dataFrame[, c("STFIPS", "LIVARAG", "SERVSETD", "REASON",
                        "LOS", "PSOURCE")]
  data$LIVARAG <- as.factor(data$LIVARAG)
  data$SERVSETD <- as.factor(data$SERVSETD)
  data$REASON <- as.factor(data$REASON)
  data$LOS <- as.factor(data$LOS)
  data$PSOURCE <- as.factor(data$PSOURCE)
  data <- data[complete.cases(data),]
  dataMatrix <- model.matrix(STFIPS~.,data)[,-1]
  dataDF <- data.frame(dataMatrix)
  dataDF$state <- data$STFIPS
  dataDF$discharge <- 1
  summedFin <- dataDF %>% group_by(state) %>% summarise_all(funs(sum))
  return(summedFin)
}

TEDSDList2014 <- c("AL", "AK", "AZ", "AR", "CA", "CO", "CT", "DE", "DC",
                     "HI", "ID", "IL", "IN", "IA","KY", "LA", "ME", "MD", "MA",
                     "MI", "MN", "MO", "MT", "NE", "NV", "NH", "NJ", "NY",
                     "NC", "ND", "OH", "OK", "OR", "PA", "RI", "SC", "SD", "TN", "TX",
                     "UT", "VT", "VA", "WA", "WI", "WY")
TEDSDList2012 <- c("AL", "AK", "AZ", "AR", "CA", "CO", "CT", "DE", "DC", "FL", "GA",
                   "HI", "ID", "IL", "IN", "IA","KY", "LA", "ME", "MD", "MA",
                   "MI", "MN", "MO", "MT", "NE", "NV", "NH", "NJ", "NY",
                   "NC", "ND", "OH", "OK", "OR", "PA", "RI", "SC", "SD", "TN", "TX",
                   "UT", "VT", "VA", "WA", "WV", "WI", "WY")

TEDSDList2013 <- c("AL", "AK", "AZ", "AR", "CA", "CO", "CT", "DE", "DC", "GA",
                   "HI", "ID", "IL", "IN", "IA", "KS","KY", "LA", "ME", "MD", "MA",
                   "MI", "MN", "MO", "MT", "NE", "NV", "NH", "NJ", "NY",
                   "NC", "ND", "OH", "OK", "OR", "PA", "RI", "SC", "SD", "TN", "TX",
                   "UT", "VT", "VA", "WA", "WV", "WI", "WY")
#death statistics

deathData20132014 <- read.csv("/Users/lex/Downloads/drug_poisoning_deaths_by_state-_us_2013_2014-v7.csv")
deathData2013 <- deathData20132014[,c("State", "X2013Number")]
deathData2013$state <- deathData2013$State
deathData2013$number <- deathData2013$X2013Number
deathData2013 <- deathData2013[,-1]
deathData2014 <- deathData20132014[,c("State", "X2014Number")]
deathData2014$state <- deathData2014$State
deathData2014 <- deathData2014[,-1]
deathData2014$number <- deathData2014$X2014Number
deathData2015 <- read.csv("/Users/lex/Downloads/2015-death-data.csv")
deathData2015$state <- deathData2015$State
deathData2015 <- deathData2015[,-1]
deathData2015$number <- deathData2015$Number
deathData2012 <- read.csv("/Users/lex/Downloads/deathdata2012.csv")


tedsD2012 <- createTEDSD(teds_d_2012)
tedsD2013 <- createTEDSD(teds_d_2013)
tedsD2014 <- createTEDSD(teds_d_2014)
tedsD2012 <- tedsD2012[-49,]
tedsD2013 <- tedsD2013[-49,]
tedsD2014 <- tedsD2014[-46,]
tedsD2014$state <- TEDSDList2014
tedsD2014 <- merge(tedsD2014, deathData2014)
tedsD2013$state <- TEDSDList2013
tedsD2013 <- merge(tedsD2013, deathData2013)
tedsD2012$state <- TEDSDList2012
tedsD2012 <- merge(tedsD2012, deathData2012)
tedsD2013$deaths <- tedsD2013$number
tedsD2012$deaths <- tedsD2012$number
tedsD2014$deaths <- tedsD2014$number
tedsD2013 <- tedsD2013[, -c(64, 65)]
tedsD2013 <- merge(census2013, tedsD2013)
tedsD2012 <- tedsD2012[, -c(65)]
tedsD2012 <- merge(census2012, tedsD2012)
tedsD2014 <- tedsD2014[, -c(65,66)]
tedsD2014 <- merge(census2014, tedsD2014)



```

```{r}




year2012 <- merge(census2012, deathData2012)
year2012 <- merge(year2012, tedsA2012)
year2012 <- merge(year2012, nsafd2012)
year2012$year <- "2012"

year2013 <- merge(census2013, deathData2013)
year2013 <- merge(year2013, tedsA2013)
year2013 <- merge(year2013, nsafd2013)
year2013$year <- "2013"

year2014 <- merge(census2014, deathData2014)
year2014 <- merge(year2014, tedsA2014)
year2014 <- merge(year2014, nsafd2014)
year2014$year <- "2014"

year2015 <- merge(census2015, deathData2015)
year2015 <- merge(year2015, tedsA2015)
year2015 <- merge(year2015, nsafd2015)
year2015$year <- "2015"


```
```{r}
library(rpart)

```
```{r}
common_cols <- intersect(colnames(year2012), colnames(year2013))
test <- rbind(
  subset(year2012, select = common_cols), 
  subset(year2013, select = common_cols)
)

common_cols2 <- intersect(colnames(test), colnames(year2014))
final1 <- rbind(
  subset(test, select = common_cols2), 
  subset(year2014, select = common_cols2)
)
```
```{r}
##create data for modeling

final <- final1

final$number <- as.numeric(gsub(",", "", final$number))
final <- final[,-(ncol(final))] 

final$publicAssistance <- final$publicAssistance/final$population
final$number <- final$number/final$population

for (column.name in names(final) [28:57]) {
  final[column.name] = final[column.name] / final$admit
}

final$admit <- final$admit/final$population

for (column.name in names(final) [59:ncol(final)]) {
  final[column.name] = final[column.name] / final$availableService
}

final$availableService <- final$availableService/final$population


final <- final[,-2]
```
```{r}

dfForTree <- final

dfScale <- final

dfScale <- dfScale[ , -which(names(dfScale) %in% c("population", "LANG1", "LANG161",
                                                   "LANG11", "LANG21", "LANG31", "LANG41", 
                                                   "LANG51", "LANG61", "LANG71", "LANG81",
                                                   "LANG91", "LANG101", "LANG121", "LANG141",
                                                   "LANG131", "LANG151", "LANG161", "LANG171",
                                                   "LANG181", "LANG191", "LANG201"))]

dfScale[,2:ncol(dfScale)] <- scale(dfScale[,2:ncol(dfScale)])

library("devtools")
#install_github("kassambara/factoextra")
library("factoextra")
```
```{r}

##pc comp

pcAll <- princomp(dfScale[,2:ncol(dfScale)], cor=TRUE, scores=TRUE)
fviz_pca_biplot(pcAll)

pcCensus <- princomp(dfScale[,2:24], cor=TRUE, scores=TRUE)
fviz_pca_biplot(pcCensus)

```
```{r}

##ML REGRESSION
library(MASS)
set.seed(1234)
dfScaleSamp <- dfScale[sample(nrow(dfScale)),]
dfTrain <- dfScaleSamp[1:122,]
dfTest <- dfScaleSamp[123:152,]

modelDF <- dfTrain
modelDF$number <- scale(modelDF$number)
ownership <- lm(number ~ OWNERSHP2 + OWNERSHP3 +
                  OWNERSHP4 + OWNERSHP5 + OWNERSHP6, dfScale[,-1])
model <- lm(number ~., dfScale[,-1])
fit2 <- stepAIC(model,direction="both")



modelResid <- resid(fit2)

plot(dfScale$number, modelResid, 
     ylab="Residuals", xlab="Level", 
     main="Level of Drug Overdose Residuals") 
abline(0, 0)

RMSE <- function(error) { sqrt(mean(error^2)) }

predictScaledDeath <- predict(fit2, dfTest)
rmseML <- sqrt(mean((predictScaledDeath - dfTest$number)^2))
```
```{r}
##CART + RANDOM FOREST
dfForTree$number <- scale(dfForTree$number)

w <- ifelse(dfForTree$number <= -1.5, "extremely low",
            ifelse((dfForTree$number >= -1.5) & (dfForTree$number <= 0), "low/medium",
                   ifelse((dfForTree$number >= 0) & (dfForTree$number <= 1.5), "medium/high",
                          "high")))
dfForTree$deathBucket <- w

x <- dfForTree[ , -which(names(dfForTree) %in% c("state", "number", "population", "REVCHK11"))]

library(rpart)
library(rpart.plot)
library(randomForest)
set.seed(4543)
x$deathBucket <- factor(x$deathBucket) 
db.rf <- randomForest(deathBucket ~ ., data=x, ntree=1000, keep.forest=FALSE,
                          importance=TRUE)
varImpPlot(db.rf)

fitDT1<-rpart(deathBucket~.,data=x)

fitDT2<-rpart(deathBucket~.,data=x,control=rpart.control(cp=0.001),xval=20) 

fitDT3<-rpart(deathBucket~.,data=x,                   
             parms=list(split="information"),   #can change to information gain
             control=rpart.control(xval=20)  )
censusX <- x[, 1:23]
censusX$deathBucket <- x$deathBucket
fitCensus <- rpart(deathBucket~.,data=censusX)
```
```{r}
##Low vs. High Death Rate States Exploartory Analysis

final1$number <- as.numeric(gsub(",", "", final1$number))
final1 <- final1[,-(ncol(final1))] 

final1$publicAssistance <- final1$publicAssistance/final1$population
final1$number <- final1$number/final1$population

for (column.name in names(final1) [28:57]) {
  final1[column.name] = final1[column.name] / final1$admit
}

final1$admit <- final1$admit/final1$population

for (column.name in names(final1) [59:ncol(final1)]) {
  final1[column.name] = final1[column.name] / final1$availableService
}

final1$availableService <- final1$availableService/final1$population


final1 <- final1[,-2]

finalCensus <- final1[,2:26]
finalCensus <- finalCensus[,-22]
finalCensus <- scale(finalCensus)

pc <- princomp(finalCensus, cor=TRUE, scores=TRUE)
fviz_pca_biplot(pc)
```
```{r}
##exploratory analysis ggplots

exp <- final1
exp$number <- scale(exp$number)
w <- ifelse(exp$number <= -1, "low", ifelse((exp$number >= -1) & (exp$number <= 1), "medium",
                                        "high"))
exp$number <- w
exp <- exp[exp$number != "medium",]
ggplot(exp, aes(publicAssistance, colour = number)) + geom_density()
ggplot(exp, aes(unemployed, colour = number)) + geom_density()
ggplot(exp, aes(publicAssistance, colour = number)) + geom_density()
ggplot(exp, aes(hsNoDiploma, colour = number)) + geom_density()
ggplot(exp, aes(medianIncome, colour = number)) + geom_density()
ggplot(exp, aes(percent_black, colour = number)) + geom_density()
ggplot(exp, aes(SUB13, colour = number)) + geom_density()
ggplot(exp, aes(SUB15, colour = number)) + geom_density()
ggplot(exp, aes(SUB114, colour = number)) + geom_density()
ggplot(exp, aes(LOC51, colour = number)) + geom_density()
ggplot(exp, aes(CTYPEHI21, colour = number)) + geom_density()
ggplot(exp, aes(FEESCALE1, colour = number)) + geom_density()
ggplot(exp, aes(SRVC1161, colour = number)) + geom_density()
ggplot(exp, aes(admit, colour = number)) + geom_density()
ggplot(exp, aes(availableService, colour = number)) + geom_density()
ggplot(exp, aes(OWNERSHP3, colour = number)) + geom_density()

#USE
ggplot(exp, aes(DETOX1, colour = number)) + geom_density()
ggplot(exp, aes(LOC51, colour = number)) + geom_density()
ggplot(exp, aes(SRVC711, colour = number)) + geom_density()
ggplot(exp, aes(CTYPE41, colour = number)) + geom_density()
ggplot(exp, aes(CTYPEHI11, colour = number)) + geom_density()
ggplot(exp, aes(CTYPEHI21, colour = number)) + geom_density()
ggplot(exp, aes(CTYPE71, colour = number)) + geom_density()
ggplot(exp, aes(CTYPE21, colour = number)) + geom_density()
```
```{r}
##Various analysis (not used in project)
subs <- final1
subs$number <- scale(subs$number)
w <- ifelse(subs$number <= -1, "low", ifelse((subs$number >= -1) & (subs$number <= 1), "medium",
                                            "high"))
subs$number <- w

high <- subs[subs$number == "high",3:140]
summedHi <- high %>% group_by(number) %>% summarise_all(funs(sum))
summedHI <- summedHi[,27:56]

for (column.name in names(summedHI) [1:29]) {
  summedHI[column.name] = summedHI[column.name] / summedHI$admit
}
library(reshape2)
nopriHi <- summedHI[,1:5]
noPriHi = melt(nopriHi)
noPriHi = noPriHi[order(noPriHi$variable),] 
library(plotly)


subHi <- summedHI[,6:24]
subHi = melt(subHi)
subHi = subHi[order(subHi$variable),]
freqHi <- summedHI[,25:29]
freqHi = melt(freqHi)
freqHi = freqHi[order(freqHi$variable),]



low <- subs[subs$number == "low",3:140]
summedLow <- low %>% group_by(number) %>% summarise_all(funs(sum))
summedLOW <- summedLow[,27:56]

for (column.name in names(summedLOW) [1:29]) {
  summedLOW[column.name] = summedLOW[column.name] / summedLOW$admit
}

nopriLOW <- summedLOW[,1:5]
nopriLOW = melt(nopriLOW)
nopriLOW = nopriLOW[order(nopriLOW$variable),] 

subLOW <- summedLOW[,6:24]
subLOW = melt(subLOW)
subLOW = subLOW[order(subLOW$variable),] 

freqLOW <- summedLOW[,25:29]
freqLOW = melt(freqLOW)
freqLOW = freqLOW[order(freqLOW$variable),] 

prior1 <- c("None", "Alcohol", "Cocaine",
           "Marijuana", "Heroin", "Methadone",
           "Other Opiates")
HI <- subHi[1:7,]$value
LOW <- subLOW[1:7,]$value
data <- data.frame(prior1, HI, LOW)

z <- plot_ly(data, x = ~prior1, y = ~HI, type = 'bar', name = 'High Rates') %>%
  add_trace(y = ~LOW, name = 'Low Rates') %>%
  layout(yaxis = list(title = 'Count'), barmode = 'group')

prior1 <- c("None in Past Month",
            "1-3 in Past Month",
            "1-2 in Past Week",
            "3-6 in Past Week", 
            "Daily")
HI <- freqHi$value
LOW <- freqLOW$value
data <- data.frame(prior1, HI, LOW)

z <- plot_ly(data, x = ~prior1, y = ~HI, type = 'bar', name = 'High Rates') %>%
  add_trace(y = ~LOW, name = 'Low Rates') %>%
  layout(yaxis = list(title = 'Count'), barmode = 'group')


```
```{r}
#2015 NSSATS analysis
nssats2015 <- merge(census2015, deathData2015)
nssats2015 <- merge(nssats2015, nsafd2015)
nssats2015 <- nssats2015[,-c(27, 28, 29)]

nssats2015$number <- as.numeric(gsub(",", "", nssats2015$number))

for (column.name in names(nssats2015) [27:ncol(nssats2015)]) {
  nssats2015[column.name] = nssats2015[column.name] / nssats2015$population
}

nssats2015$publicAssistance <- nssats2015$publicAssistance/nssats2015$population

nsScale <- nssats2015[,3:27]
nsScale <- scale(nsScale)
nsScale <- nsScale[ , -22]
row.names(nsScale) <- nssats2015[,1]

pc <- princomp(nsScale, cor=TRUE, scores=TRUE)
fviz_pca_biplot(pc)


nssats2015 <- nssats2015[,-2]

nssats2015$number <- scale(nssats2015$number)

w <- ifelse(nssats2015$number <= -1.0, "extremely low",
            ifelse((nssats2015$number >= -1.0) & (nssats2015$number <= 0), "low/medium",
                   ifelse((nssats2015$number >= 0) & (nssats2015$number <= 1.0), "medium/high",
                          "high")))
nssats2015$deathBucket <- w

x <- nssats2015[ , -which(names(nssats2015) %in% c("state", "number", "population"))]
##CART for 2015
library(rpart)
fitDT<-rpart(deathBucket~.,data=x)

fitDT<-rpart(deathBucket~.,data=x,control=rpart.control(cp=0.001),xval=20) 

fitDT<-rpart(deathBucket~.,data=x,                   
             parms=list(split="information"),   #can change to information gain
             control=rpart.control(xval=20)  )






